Quantifying maternal reproductive output of chondrichthyan fishes

### 2024-06-23 reports/main-text.tex
repo_prep

For the live-bearing and egg-laying class of chondrichthyan fishes a three parameter logistic `maternity' function with a variable upper asymptote, \(P_{Max}\), can be used to predict the average probability of a female giving birth or laying eggs in a season. Although fundamental to calculating the reproductive capacity of a population, few studies report maternity functions, with maturity functions often used as a proxy. Applying logistic models to simulated and empirical data showed that it was feasible to estimate \(P_{Max}\) from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used. Maturity parameters could also be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output. Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models. In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction. This study suggests that for sample sizes \textgreater{} 100, estimation of \(P_{Max}\) is generally preferable to using a fixed value.

@@

Walker (2005) recognised the need for a maternity function, as distinct from a maturity function, to accurately quantify reproductive capacity in chondrichthyans. For this purpose he used a non-linear, three-parameter logistic function (3PLF) of the form: \[Y_i \sim B(x_i; n_i, P_i)\] \[E[Y_i] = n_iP_i = \frac{\alpha}{1 + e^{-(b_0 + b_1x_i)}}\] where the proportion of individuals in a given length or age category, \(x_i\), that are in maternal condition, \(Y_i\), is a binomially distributed random variable, with an expected value equal to the probability of an individual from that category being in maternal condition, \(P_i\), multiplied by the total number of individuals in the category, \(n_i\). If \(x\) is a continuous rather than categorical variable (i.e.~\(n = 1\)), \(Y_i\) reduces to a Bernoulli random variable (Zuur et al. 2007). \(\alpha\) is the upper asymptote of the curve, controlling the maximum value of \(Y\) as \(x\) approaches infinity. The lesser-used 3PLF reduces to the ubiquitous 2PLF when \(\alpha = 1\), as is appropriate for most teleost and chondrichthyan populations if modeling maturity (Quinn and Deriso 1999, Roa et al. 1999). When \(\alpha = 1\), the function also belongs to the class of generalised linear models; logistic transformation of the dependent variable allows \(Y\) to be modeled as a linear function of \(x_i\), where \(b_0\) and \(b_1\) are the intercept and slope, respectively (Zuur et al. 2007). Alternatively, the model can be expressed in more biologically relevant terms as: \[E[Y_i] = P_{Max} \left ( 1 + e^{-ln(19)\left ( \frac{x_i - x_{50}}{x_{95} - x_{50}} \right )}  \right )^{-1} \] where \(x_{50}\) and \(x_{95}\) are the sizes or ages at which 50\% and 95\% of the maximum proportion of individuals (\(P_{Max}\)) are in maternal condition. Instead of being fit to data on female maturity condition the model is fit to dichotomous data on female maternal condition (non maternal / maternal).

@@

Wider use of maternity functions in studies of chondrichthyan reproductive biology is needed to improve calculation of reproductive capacity in the quantitative population assessments that are increasingly being used to support fisheries management and prioritize conservation actions (Cortés et al. 2012). It may also help address the lack of empirical data on the periodicity of chondrichthyan reproductive cycles. Historically, reproductive periodicity has only been determined qualitatively, and has been assumed to be fixed (e.g.~annual, biennial), despite observations of plasticity in this trait within discrete populations (Higgs et al. 2020). Information on reproductive periodicity is contained within maternal data, however the existing approach of fixing \(P_{Max}\) in maternity functions prevents statistical inference from being made on these data. Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive periodicity. To date only a single study on spiny dogfish, \emph{Squalus acanthias}, has statistically estimated \(P_{Max}\) from data (Colonello et al. 2016), and the feasibility of doing so for a wider range of species has not been investigated.

@@

\[
\hat{C}_{l} = \sum_a(F_{l}/Z_{l})(1-\exp(-Z_{l}))N_{l,a}
\]

To examine the effect of varying sample size, six scenarios were conducted with catch ranging from 50 to 2500 individuals, reflecting a gradient from data poor to rich.

@@

Parameter estimation was generally straightforward; with sample sizes \(> 50\) a convergence success rate of \(\geq 95\%\) was achieved in all scenarios (Figure S1, Table S1). Overall, convergence and parameter boundary failures were most common with the 2PLF-maternity method which essentially involved fitting data to an under-parameterized model where \(P_{Max}\) was always fixed at 1. In addition to convergence failures, 12 simulations also generated insufficient maternal data (zero or one maternal females in the simulated data), and were discarded.

Generally when using the 3PLF methods model performance declined as the periodicity of the reproductive cycle increased. The nature of this problem can be seen in the positive correlation between the \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\) parameters whereby \(P_{Max}\) affects the magnitude and direction of bias in these parameters (Figure 2). With lower values of \(P_{Max}\) it was apparently more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote. As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\hat{P_{Max}}\) (Figure 2).

For the 3PLF methods, parameter estimates tended to be biased high at lower sample sizes (Figure S2, Figure S3). In some scenarios, particularly with sample sizes \(\leq\) 250, fixing the asymptote was effective at reducing bias, although for larger sample sizes neither method was clearly preferable (Figure S2, Figure S3). The accuracy of parameter estimates varied considerably across different combinations of variables used in the simulation and among parameters (Figure S4, Figure S5). Again, in some scenarios with smaller sample sizes \(\leq\) 250, fixing the asymptote improved accuracy with the 3PLF methods, although at larger sample sizes better accuracy was achieved by estimating \(P_{Max}\) (Figure 3). Overall, \(P_{Max}\) was comparatively more challenging to estimate accurately than \(L'_{50}\) (Figure S4, Figure S5).

While the 3PLF-fixed method was in some cases able to reduce bias at lower sample sizes, a trade off in the use of this method was poorer interval coverage (Figure 4, Figure S6). For most simulations, interval coverage for \(\hat{L'_{50}}\) was well below the expected level of 50\%, irrespective of sample size. In contrast, for the 3PLF-estimated method interval coverage oscillated around 50\% for both \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\).

Differences in life history and gear selectivity played an important role in the ability to obtain precise and unbiased maternity parameters. Certain combinations of variables in the simulated data resulted in fewer females in maternal or immature condition making parameter estimation difficult (Figure S8, Figure S9). These effects were not necessarily consistent between species and seemed to reflect differences in the underlying population length structure. For example, in low selectivity scenarios for gummy sharks \textasciitilde60\% of individuals were immature compared to \textasciitilde50\% in the corresponding scenarios for school sharks (Figure S9). As \(P_{Max}\) decreased the proportion of maternal females in the analysis also decreased.

@@

The effect of the different methods in ultimately calculating \(R_0\) varied considerably across the various scenarios. For school sharks, estimating \(P_{Max}\), was the most effective way to minimize bias in most scenarios (Figure 5, Figure S10). Substituting maturity parameters (2PLF-maturity method) led to a median overestimation of \(R_0\) by 3.4\% across all scenarios, while the 2PLF-maternity method led to a median underestimate of 11.1\%. (Table S1, Figure S10). In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in several instances (Figure 5, Figure S10). Using the 2PLF method for gummy sharks led to a median overestimate of \(R_0\) of 22\% (Figure S10). For both species, better accuracy in calculating \(R_0\) was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11). For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of \(P_{Max}\).

@@

Maternal data were re-analysed for 1087 sandbar sharks including 0 mature individuals of which \ensuremath{\infty{}}\% were in maternal condition. Using the 3PLF-estimated method, the maximum likelihood estimate for \(\hat{P_{Max}}\) was 0.48 (Table 2). Despite having a sample size of \textgreater{} 600 mature females, the proportion of maternal individuals at length was still uncertain and \(\hat{P_{Max}}\) was estimated to lie between 0.39 and 0.60 with 95\% confidence. For 3PLF-fixed models a value of \(P_{Max} = 0.5\) was strongly supported (\(\Delta_{AIC}\) = 14.15) over the alternative fixed value of \(P_{Max} = 0.33\). The model with \(P_{Max} = 0.5\) also outperformed the 3PLF-estimated method (\(\Delta_{AIC}\) = 1.86). Given the fixed model had one fewer estimated parameters, both models had essentially the same level of support given the data (Burnham and Anderson 2002).

@@

Relatively few reproductive biology studies have used maternity functions to model maternal reproductive output in chondrichthyan fishes. Where they have been employed the approach has typically been to use a three parameter logistic function with a fixed, user-defined value for the upper asymptote, \(P_{Max}\). Here we show that it is feasible to estimate \(P_{Max}\) from maternal data, in turn enabling statistical inferences on reproductive periodicity. Applying 3PLF models with estimated and fixed values of \(P_{Max}\) to simulated data showed that accuracy, bias, and confidence interval coverage often improved when \(P_{Max}\) was estimated. Using a fixed value for \(P_{Max}\) in some cases resulted in lower bias at low sample sizes. This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters. Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology. We conclude by discussing the advantages of adopting this approach to quantifying maternal reproductive output and future directions.

@@

While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately. Outcomes of the simulation study suggested that samples sizes of \(\geq500\) would be needed to estimate \(P_{Max}\) with \textless10\% MARE for a triennially reproducing species. This is a much larger error than is achievable from conventional logistic maturity analysis (Roa et al. 1999). In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying \(R_0\). Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a sample size of \textgreater{} 1000. A similar sample size was used by Colonello \emph{et al}. (2016) to estimate plausible values of \(P_{Max}\) and maternity parameters for south Atlantic spiny dogfish. These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained (Oddone et al. 2010, Tribuzio and Kruse 2012).

In light of the data requirements suggested by this study, placing a penalty function on \(P_{Max}\) or using a the use of a prior distribution

For some species, and data on pregnant females are particularly sparse,
the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans. From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging. However, the extent to which maturity parameters can provide a good approximation of maternity parameters may be species-specific, depending on how close \(L_{50}\) is to \(L'_{50}\). While several studies have shown these parameters to be similar (Walker 2007, Soto-López et al. 2018), \(L'_{50}\) is more frequently shifted to the right of \(L_{50}\) (Braccini et al. 2006, Montealegre-Quijano et al. 2014, Colonello et al. 2016, Palacios-Hernández et al. 2020), and this length difference can equate to one or more years. For example, Harry et al. (2013) found both spot-tail shark, \emph{C. sorrah}, and Australian blacktip shark, \emph{C. tilstoni}, began reproducing the year after reaching sexual maturity. Similarly, Fujinami et al. (2017) found \(A'_{50}\) to be 1.4 years older than \(A_{50}\) in blue sharks.

@@

The empirical analysis carried out using data for the sandbar shark illustrates how novel insights can be gained from maternal data by estimating \(P_{Max}\). Detailed reproductive studies of the western North Atlantic population by Baremore and Hale (2012) and Piercy et al. (2016) were ambiguous about the frequency of reproduction. In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or possibly longer at the population level. Reanalysis of these combined data sets using the 3PLF-estimated method, however, resulted in a maximum likelihood estimate of \(\hat{P_{Max}} = 0.48\), considerably larger than 0.37 - the proportion of pregnant mature females observed by Baremore and Hale (2012). Similarly, based on AIC values, a fixed value of \(P_{Max} = 0.5\) had a much higher relative plausibility than a value of 0.33 (Table 2).

This unexpected result may be related to the gradual attainment of asymptotic maternal status. Maturity in sandbar sharks appears to occur over an extended size range. \(L_{5}\) and \(L_{95}\) are 146cm and 176cm -- a length interval that corresponds to around seven years given the sandbar shark's slow growth rate (Hale and Baremore 2013). With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently. The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially, although few females appear to attain this size (Hale and Baremore 2013). Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund (Baremore and Hale 2012), they also reproduce more frequently. This implies that size selective fishing or the removal of larger and older female sharks could have a disproportionately greater impact on population productivity.

@@

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-baremore_reproduction_2012}
Baremore, I.E., and Hale, L.F. 2012. Reproduction of the {Sandbar} {Shark} in the {Western} {North} {Atlantic} {Ocean} and {Gulf} of {Mexico}. Marine and Coastal Fisheries \textbf{4}(1): 560--572. doi: \href{https://doi.org/10.1080/19425120.2012.700904}{10.1080/19425120.2012.700904}.

\bibitem[\citeproctext]{ref-baremore_reproduction_2013}
Baremore, I.E., and Passerotti, M.S. 2013. Reproduction of the {Blacktip} {Shark} in the {Gulf} of {Mexico}. Marine and Coastal Fisheries \textbf{5}(1): 127--138. doi: \href{https://doi.org/10.1080/19425120.2012.758204}{10.1080/19425120.2012.758204}.

\bibitem[\citeproctext]{ref-bowlby_implications_2020}
Bowlby, H.D., and Gibson, A.J.F. 2020. Implications of life history uncertainty when evaluating status in the {Northwest} {Atlantic} population of white shark (\emph{{Carcharodon} carcharias)}. Ecology and Evolution \textbf{10}(11): 4990--5000. doi: \href{https://doi.org/10.1002/ece3.6252}{10.1002/ece3.6252}.

@@

\bibitem[\citeproctext]{ref-burnham_model_2002}
Burnham, K.P., and Anderson, D.R. 2002. Model selection and multimodel inference: A practical information-theoretic approach. 2nd edn. Springer, New York, NY.

@@

\bibitem[\citeproctext]{ref-dulvy_you_2008}
Dulvy, N.K., Baum, J.K., Clarke, S., Compagno, L.J.V., Cortes, E., Domingo, A., Fordham, S., Fowler, S., Francis, M.P., Gibson, C., Martínez, J., Musick, J.A., Soldo, A., Stevens, J.D., and Valenti, S. 2008. You can swim but you can't hide: The global status and conservation of oceanic pelagic sharks and rays. Aquatic Conservation: Marine and Freshwater Ecosystems.

@@

\bibitem[\citeproctext]{ref-frisk_life_2005}
Frisk, M.G., Miller, T.J., and Dulvy, N.K. 2005. Life histories and vulnerability to exploitation of elasmobranchs: {Inferences} from elasticity, pertubation and phylogenetic analyses. Journal of Northwest Atlantic Fishery Science \textbf{35}: 27--45. doi: \href{https://doi.org/10.2960/J.v35.m514}{10.2960/J.v35.m514}.

\bibitem[\citeproctext]{ref-fujinami_reproductive_2017}
Fujinami, Y., Semba, Y., Okamoto, H., Ohshimo, S., and Tanaka, S. 2017. Reproductive biology of the blue shark (\emph{{Prionace} glauca}) in the western {North} {Pacific} {Ocean}. Mar. Freshwater Res. \textbf{68}(11): 2018--2027. Available from \url{https://doi.org/10.1071/MF16101}.

@@

\bibitem[\citeproctext]{ref-hale_age_2013}
Hale, L.F., and Baremore, I.E. 2013. Age and {Growth} of the {Sandbar} {Shark} (\emph{{Carcharhinus} plumbeus}) from the {Northern} {Gulf} of {Mexico} and the {Western} {North} {Atlantic} {Ocean}. Gulf of Mexico Science \textbf{1}(2): 28--39.

@@

\bibitem[\citeproctext]{ref-jennings_marine_2001}
Jennings, S., Kaiser, M., and Reynolds, J.D. 2001. Marine fisheries ecology. John Wiley \& Sons, Oxford.

@@

\bibitem[\citeproctext]{ref-king_fisheries_2007}
King, M. 2007. Fisheries biology, assessment and management. Blackwell Publishing, Oxford.

>>>>

For the live-bearing and egg-laying class of chondrichthyan fishes a three parameter logistic `maternity' function with a variable upper asymptote, \(P_{Max}\), can be used to predict the average probability of a female giving birth or laying eggs in a season. Although fundamental to calculating the reproductive capacity of a population, few studies report maternity functions, with maturity functions often used as a proxy. Applying logistic models to simulated and empirical data showed that it was feasible to estimate \(P_{Max}\) from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used. However, sample sizes of 100 - 200 maternal females were typically required for accurate estimation of \(P_{Max}\). While maturity parameters could be estimated with greater accuracy, substituting them for maternity parameters overestimated lifetime reproductive output. Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models. In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction.

@@

Walker (2005) recognised the need for a maternity function, as distinct from a maturity function, to accurately quantify reproductive capacity in chondrichthyans. For this purpose he used a non-linear, three-parameter logistic function (3PLF) of the form: \[Y_i \sim B(x_i; n_i, P_i)\] \[E[Y_i] = n_iP_i = \frac{\alpha}{1 + e^{-(b_0 + b_1x_i)}}\] where the proportion of individuals in a given length or age category, \(x_i\), that are in maternal condition, \(Y_i\), is a binomially distributed random variable, with an expected value equal to the probability of an individual from that category being in maternal condition, \(P_i\), multiplied by the total number of individuals in the category, \(n_i\). If \(x\) is a continuous rather than categorical variable (i.e.~\(n = 1\)), \(Y_i\) reduces to a Bernoulli random variable (Zuur et al. 2007). The upper asymptote of the curve, \(\alpha\), controls the maximum value of \(Y\) as \(x\) approaches infinity. The lesser-used 3PLF reduces to the ubiquitous 2PLF when \(\alpha = 1\), as is appropriate for most teleost and chondrichthyan populations if modeling maturity (Quinn and Deriso 1999, Roa et al. 1999). When \(\alpha = 1\), the function also belongs to the class of generalised linear models; logistic transformation of the dependent variable allows \(Y\) to be modeled as a linear function of \(x_i\), where \(b_0\) and \(b_1\) are the intercept and slope, respectively (Zuur et al. 2007). Alternatively, the model can be expressed in more biologically relevant terms as: \[E[Y_i] = P_{Max} \left ( 1 + e^{-ln(19)\left ( \frac{x_i - x_{50}}{x_{95} - x_{50}} \right )}  \right )^{-1} \] where \(x_{50}\) and \(x_{95}\) are the sizes or ages at which 50\% and 95\% of the maximum proportion of individuals (\(P_{Max}\)) are in maternal condition. Instead of being fit to data on female maturity condition the model is fit to dichotomous data on female maternal condition (non maternal / maternal).

@@

Wider use of maternity functions in studies of chondrichthyan reproductive biology is needed to improve calculation of reproductive capacity in the quantitative population assessments that are increasingly being used to support fisheries management and prioritize conservation actions (Cortés et al. 2012, Swenson et al. 2024). It may also help address the lack of empirical data on the periodicity of chondrichthyan reproductive cycles. Historically, reproductive periodicity has only been determined qualitatively, and has been assumed to be fixed (e.g.~annual, biennial), despite observations of plasticity in this trait within discrete populations (Higgs et al. 2020, Nosal et al. 2021). Information on reproductive periodicity is contained within maternal data, however the existing approach of fixing \(P_{Max}\) in maternity functions prevents statistical inference from being made on these data. Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive cycles. To date only a single study on spiny dogfish, \emph{Squalus acanthias}, has statistically estimated \(P_{Max}\) from data (Colonello et al. 2016), and the feasibility of doing so for a wider range of species has not been investigated.

@@

\[
C_{l} = \sum_a(F_{l}/Z_{l})(1-\exp(-Z_{l}))N_{l,a}
\]

To examine the effect of varying sample size, six scenarios were conducted with catch ranging from 50 to 2500 individuals, reflecting a gradient from data poor to rich. Length data were generated by randomly sampling, with replacement, from a vector of length classes, weighted by a vector of probabilities determined from the above catch equation.

@@

Parameter estimation was generally straightforward; with sample sizes \(> 50\) a convergence success rate of \(\geq 95\%\) was achieved in all scenarios (Figure S1, Table S1). Overall, convergence and parameter boundary failures were most common with the 2PLF-maternity method which essentially involved fitting data to an under-parameterized model where \(P_{Max}\) was always fixed at 1. In addition to convergence failures, 12 simulations also generated zero maternal females, and were discarded.

When using the 3PLF methods model performance typically declined as the periodicity of the reproductive cycle increased. This is seen in the positive correlation between \(\widehat{L'_{50}}\) and \(\widehat{P_{Max}}\), whereby true underlying \(P_{Max}\) affected the magnitude and direction of bias in these parameters (Figure 2). With lower values of \(P_{Max}\) it appeared to be more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote. As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\widehat{P_{Max}}\). Additionally, longer reproductive cycles resulted in fewer females in maternal condition, which was associated with larger bias (Figure 2).

For the 3PLF methods, parameter estimates tended to be biased high at lower sample sizes (Figure S2, Figure S3). In some scenarios, particularly with sample sizes \(\leq\) 250, fixing the asymptote was effective at reducing bias, although for larger sample sizes neither method was clearly preferable (Figure S2, Figure S3). The accuracy of parameter estimates varied considerably across different combinations of variables used and among parameters (Figure S4, Figure S5). Again, in some scenarios with smaller sample sizes \(\leq\) 250, fixing the asymptote improved accuracy with the 3PLF methods, although at larger sample sizes better accuracy was achieved by estimating \(P_{Max}\) (Figure 3). Overall, \(P_{Max}\) was comparatively more challenging to estimate accurately than \(L'_{50}\) (Figure S4, Figure S5).

While the 3PLF-fixed method was in some cases able to reduce bias at lower sample sizes, a trade off in the use of this method was poorer interval coverage (Figure 4, Figure S6). For most simulations, interval coverage for \(\widehat{L'_{50}}\) was well below the expected level of 50\%, irrespective of sample size. In contrast, for the 3PLF-estimated method interval coverage oscillated around 50\% for both \(\widehat{L'_{50}}\) and \(\widehat{P_{Max}}\).

Differences in life history and gear selectivity played an important role in the ability to obtain accurate and unbiased maternity parameters. Certain combinations of variables in the simulated data resulted in few females in maternal or immature condition, making parameter estimation difficult (Figure S8, Figure S9). This was particularly the case for \(P_{Max}\) which required a relatively high maternal sample size to enable accurate parameter estimation (Figure 5). These effects were not necessarily consistent between species and seemed to reflect differences in the underlying population length structure. For example, in biennial low selectivity scenarios for gummy sharks \textasciitilde60\% of individuals were immature compared to \textasciitilde50\% in the corresponding scenarios for school sharks (Figure S9).

@@

The effect of the different methods in ultimately calculating \(R_0\) varied considerably across the various scenarios. For school sharks, estimating \(P_{Max}\), was the most effective way to minimize bias in most scenarios (Figure S10). Substituting maturity parameters (2PLF-maturity method) led to a median overestimation of \(R_0\) by 3.4\% across all scenarios, while the 2PLF-maternity method led to a median underestimate of 11.1\%. (Figure S10). In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in many instances (Figure S11). Using the 2PLF method for gummy sharks led to a median overestimate of \(R_0\) of 22\% (Figure S10). For both species, better accuracy in calculating \(R_0\) was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S12). For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of \(P_{Max}\).

@@

Maternal data were re-analysed for 1087 sandbar sharks including 640 mature individuals of which 32\% were in maternal condition. Using the 3PLF-estimated method, the maximum likelihood estimate for \(\widehat{P_{Max}}\) was 0.48 (Table 2). Despite having a sample size of \textgreater{} 600 mature females, the proportion of maternal individuals at length was still uncertain and \(\widehat{P_{Max}}\) was estimated to lie between 0.39 and 0.60 with 95\% confidence (Figure 7). For 3PLF-fixed models a value of \(P_{Max} = 0.5\) was strongly supported (\(\Delta_{AIC}\) = 14.15) over the alternative fixed value of \(P_{Max} = 0.33\). The model with \(P_{Max} = 0.5\) also outperformed the 3PLF-estimated method (\(\Delta_{AIC}\) = 1.86). Given the fixed model had one fewer estimated parameter, both models had essentially the same level of support given the data (Burnham and Anderson 2002).

@@

Relatively few reproductive biology studies have used maternity functions to model maternal reproductive output in chondrichthyan fishes. Where they have been employed the approach has typically been to use a three parameter logistic function with a fixed, user-defined value for the upper asymptote, \(P_{Max}\). Here we show that it is feasible to estimate \(P_{Max}\) from maternal data, in turn enabling statistical inferences on reproductive periodicity. Applying 3PLF models with estimated and fixed values of \(P_{Max}\) to simulated data showed that accuracy, bias, and confidence interval coverage often improved when \(P_{Max}\) was estimated. Using a fixed value for \(P_{Max}\) in some cases reduced bias. Overall, however, relatively large sample sizes were needed to accurately estimate \(P_{Max}\), particularly with longer reproductive cycles. This study also demonstrated that, in some cases, recruitment was overestimated when maturity data were used to approximate maternal data and inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters. Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology. We conclude by discussing the advantages of adopting this approach and future directions.

@@

Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate \(P_{Max}\), but was difficult to do so accurately. For non-annual reproductive cycles, \(> 100\) school sharks and \(> 200\) gummy sharks in maternal condition were typically needed to estimate \(P_{Max}\) with \textless10\% MARE (Figure 5). This indicates that the both sample size requirements and levels of uncertainty are likely to be larger than conventional logistic maturity analysis (Roa et al. 1999). In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in minimizing bias in \(R_0\) (Figure S11). Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a total sample size of \textgreater{} 1000 females, 206 of which were in maternal condition. A similar sample size was used by Colonello \emph{et al}. (2016) to estimate plausible values of \(P_{Max}\) and maternity parameters for south Atlantic spiny dogfish.

The simulation study also also identified some situations where it may be preferable to fix \(P_{Max}\), particularly at lower sample sizes (\(\le\) 100 individuals total). Providing clear guidelines on this is difficult as model performance was species-specific and influenced by life history and gear selectivity. The success of any analysis using a fixed value also ultimately depends on the analyst choosing a suitable value. The potential to introduce a greater level of bias through the incorrect selection of \(P_{Max}\) should also be considered if using this approach (Natanson et al. 2019).

Together, these results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained (Oddone et al. 2010, Tribuzio and Kruse 2012). For smaller samples sizes, placing a prior or penalty on \(P_{Max}\) may also help overcome issues of parameter inaccuracy, enabling existing information about reproductive periodicity to be incorporated from other populations or closely-related species (Smart and Grammer 2021). Information on length at maturity, which is likely to be relatively easier to estimate, may also be informative in the development of priors, with \(L_{50}\) effectively providing a lower bound for \(L'_{50}\).

For data-poor chondrichthyan species, or where data on pregnant females are sparse or unattainable, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option. From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging. However, the extent to which maturity parameters can provide a good approximation of maternity parameters may be species-specific, depending on how close \(L_{50}\) is to \(L'_{50}\). While several studies have shown these parameters to be similar (Walker 2007, Soto-López et al. 2018), \(L'_{50}\) is more frequently shifted to the right of \(L_{50}\) (Braccini et al. 2006, Montealegre-Quijano et al. 2014, Colonello et al. 2016, Palacios-Hernández et al. 2020), and this length difference can equate to one or more years. For example, Harry et al. (2013) found both spot-tail shark, \emph{C. sorrah}, and Australian blacktip shark, \emph{C. tilstoni}, began reproducing the year after reaching sexual maturity. Similarly, Fujinami et al. (2017) found \(A'_{50}\) to be 1.4 years older than \(A_{50}\) in blue sharks.

@@

The empirical analysis carried out using data for the sandbar shark illustrates how novel insights can be gained from maternal data by estimating \(P_{Max}\). Detailed reproductive studies of the western North Atlantic population by Baremore and Hale (2012) and Piercy et al. (2016) were ambiguous about the frequency of reproduction. In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or possibly longer at the population level. Reanalysis of these combined data sets using the 3PLF-estimated method, however, resulted in a maximum likelihood estimate of \(\widehat{P_{Max}} = 0.48\), considerably larger than 0.37 - the proportion of pregnant mature females observed by Baremore and Hale (2012). Similarly, based on AIC values, a fixed value of \(P_{Max} = 0.5\) had a much higher relative plausibility than a value of 0.33 (Table 2).

This unexpected result may be related to the gradual attainment of asymptotic maternal status. Maturity in sandbar sharks appears to occur over an extended size range. \(L_{5}\) and \(L_{95}\) are 146cm and 176cm -- a length interval that corresponds to around seven years given the sandbar shark's slow growth rate (Hale and Baremore 2013). With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently, while larger sharks may ultimately be capable of reproducing biennially. Given that few females appear to attain this size (Hale and Baremore 2013), the modal reproductive frequency is still likely at least triennial within the population. Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund (Baremore and Hale 2012), they also reproduce more frequently. This implies that size selective fishing or the removal of larger and older female sharks could have a disproportionately greater impact on population productivity.

@@

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-baremore_reproduction_2012}
Baremore, I.E., and Hale, L.F. 2012. Reproduction of the {Sandbar} {Shark} in the {Western} {North} {Atlantic} {Ocean} and {Gulf} of {Mexico}. Mar. Coast. Fish. \textbf{4}(1): 560--572. doi: \href{https://doi.org/10.1080/19425120.2012.700904}{10.1080/19425120.2012.700904}.

\bibitem[\citeproctext]{ref-baremore_reproduction_2013}
Baremore, I.E., and Passerotti, M.S. 2013. Reproduction of the {Blacktip} {Shark} in the {Gulf} of {Mexico}. Mar.Coast. Fish. \textbf{5}(1): 127--138. doi: \href{https://doi.org/10.1080/19425120.2012.758204}{10.1080/19425120.2012.758204}.

\bibitem[\citeproctext]{ref-bowlby_implications_2020}
Bowlby, H.D., and Gibson, A.J.F. 2020. Implications of life history uncertainty when evaluating status in the {Northwest} {Atlantic} population of white shark (\emph{{Carcharodon} carcharias)}. Eco. Evol. \textbf{10}(11): 4990--5000. doi: \href{https://doi.org/10.1002/ece3.6252}{10.1002/ece3.6252}.

@@

\bibitem[\citeproctext]{ref-burnham_model_2002}
Burnham, K.P., and Anderson, D.R. 2002. Model {Selection} and {Multimodel} {Inference}: A {Practical} {Information-theoretic} {Approach}. {Second} {Edition}. Springer, New York, NY.

@@

\bibitem[\citeproctext]{ref-dulvy_you_2008}
Dulvy, N.K., Baum, J.K., Clarke, S., Compagno, L.J.V., Cortes, E., Domingo, A., Fordham, S., Fowler, S., Francis, M.P., Gibson, C., Martínez, J., Musick, J.A., Soldo, A., Stevens, J.D., and Valenti, S. 2008. You can swim but you can't hide: The global status and conservation of oceanic pelagic sharks and rays. Aquat. Conserv. Mar. Freshw. Ecosyst.

@@

\bibitem[\citeproctext]{ref-frisk_life_2005}
Frisk, M.G., Miller, T.J., and Dulvy, N.K. 2005. Life histories and vulnerability to exploitation of elasmobranchs: {Inferences} from elasticity, pertubation and phylogenetic analyses. J. Northwest Atl. Fish. Sci. \textbf{35}: 27--45. doi: \href{https://doi.org/10.2960/J.v35.m514}{10.2960/J.v35.m514}.

\bibitem[\citeproctext]{ref-fujinami_reproductive_2017}
Fujinami, Y., Semba, Y., Okamoto, H., Ohshimo, S., and Tanaka, S. 2017. Reproductive biology of the blue shark (\emph{{Prionace} glauca}) in the western {North} {Pacific} {Ocean}. Mar. Freshwater Res. \textbf{68}(11): 2018--2027. doi: \href{https://doi.org/10.1071/MF16101}{10.1071/MF16101}.

@@

\bibitem[\citeproctext]{ref-hale_age_2013}
Hale, L.F., and Baremore, I.E. 2013. Age and {Growth} of the sandbar shark (\emph{{Carcharhinus} plumbeus}) from the {Northern} {Gulf} of {Mexico} and the {Western} {North} {Atlantic} {Ocean}. Gulf Mex. Sci. \textbf{1}(2): 28--39.

@@

\bibitem[\citeproctext]{ref-jennings_marine_2001}
Jennings, S., Kaiser, M., and Reynolds, J.D. 2001. Marine {Fisheries} {Ecology}. John Wiley \& Sons, Oxford.

@@

\bibitem[\citeproctext]{ref-king_fisheries_2007}
King, M. 2007. Fisheries {Biology}, {Assessment} and {Management}. Blackwell Publishing, Oxford.
